Kelp morphometric properties and the link to hydrodynamic modelling

### 2018-08-28 chapter_1.Rmd
updated bib. file, added hyrodynamic modelling section

Seaweeds, browns in particular, are capable of adapting morphological characteristics to persist in changing and variable ocean environments. Changes in morphology have been shown to increase photosynthetic ability, enhance nutrient uptake, and reduce drag with regards to changes in temperature and wave exposure. Wave exposure has been shown to be an important driver of seaweed morphology, as the main mechanism of seaweed mortality is through the mechanism of dislodgment. Changing morphology, such as decreasing frond area or strengthening through stipe circumference increase/reduces drag and increases the probability of survival. However, locally adapted seaweed may still be dislodged in pulse disturbance events such as storms, and may raft far distances and wash up on beaches. An example would be *Ecklonia maxima* which is a conspicuous brown seaweed along the coastline and beaches around South Africa. Therefore, the beach-cast may not always originate from adjacent kelp populations but rather from other regions which the individual is adapted for. Therefore, the kelp morphology may act as a proxy for investigating coastal currents and changes thereof. The advances in ocean hydrodynamic modelling has made great progress and has been applied in a variety of ways. Therefore, this study will use advances in hydrodynamic modelling in combination with kelp morphological characteristics to investigate coastal currents. Once the model has been established it may be applied in other ways, such as investigating the transport of micorplastics along the South African coastline. The harms of microplastics to the marine environment has gained much traction in recent years, but research in South Africa is lacking. The coupling of kelp morphology and the ability to simulate hydrodynamic processes can greatly improve our understanding of transport pathways and likely locations of accumulation. This in turn may inform management decisions with regards to eliminating and managing marine pollution in South Africa.

@@

>>>>

Seaweeds, browns in particular, are capable of adapting morphological characteristics to persist in changing and variable ocean environments. Changes in morphology have been shown to increase photosynthetic ability, enhance nutrient uptake, and reduce drag with regards to changes in temperature and wave exposure. Wave exposure has been shown to be an important driver of seaweed morphology, as the main mechanism of seaweed mortality is through the dislodgment. Changing morphology reduces drag and increases the probability of survival. However, locally adapted seaweed may still be dislodged in pulse disturbance events such as storms, and may raft far distances and wash up on beaches. An example would be *Ecklonia maxima* which is a conspicuous brown seaweed along the coastline and beaches around South Africa. Therefore, the beach-cast may not always originate from adjacent kelp populations but rather from other regions which the individual is adapted for. Therefore, the kelp morphology may act as a proxy for investigating coastal currents and changes thereof. The advances in ocean hydrodynamic modelling has made great progress and has been applied in a variety of ways. Therefore, this study will use advances in hydrodynamic modelling in combination with kelp morphological characteristics to investigate coastal currents. Once the model has been established it may be applied in other ways, such as investigating the transport of micorplastics along the South African coastline. The harms of microplastics to the marine environment has gained much traction in recent years, but research in South Africa is lacking. The coupling of kelp morphology and the ability to simulate hydrodynamic processes can greatly improve our understanding of transport pathways and likely locations of accumulation. This in turn may inform management decisions with regards to eliminating and managing marine pollution in South Africa.

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment. Like temperature, wave exposure can act on regional scales and has been shown to play a role in determing distribution, abundance, diversity, composition and productivity of benthic and rocky shore communties. For example, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake through increased turbulence in the boundry layer [@lobban1994]. @lewis found that degrees of wave exposure vertical structure rocky shore communities...**work in progress**...

# Hydrodynamic modelling

Traditional ecological measures of wave exposure usually incorportates integrative measures of hyrdodynamic conditions at a particular site. More specifically, it is the integration of mechanical processes and the influence that is has on the ecology of nearshore communities.

Wave exposure may be modelled through various methods which range from simple cartographic to more advanced numerical wave models. Cartographical models can be qualitative or quantitative and were designed for the need of wave exposure measures to explain ecological distributions. A simple set of calculations on coastline and wind data, and relatively small input data sets are required. These are regarded as "fetch-based models", which measure the length of open water associated with a site along a straight line. The output of such an approach is a simplified estimate of the potential wave energy for a specific set of sites. Advances in cartographical methods using fetch-based models has allowed for wave exposure measurments for larger areas, and has been suggested as a method for predicting macroalgal community structure [@burrows2008]. The study...**insert detailed description here**...

@lindegarth2005 critized this approach, arguing that the choice of wave exposure method  can influence ecological inference. The authors also highlighted the need for objective, reproducible and quantitative studies comparing exposure indices [@lindegarth2005]. A study by @sundbald...**include description of study here**...Other authors, such as @hill2010, have argued that these simple measures can be improved upon by including bathymetery data which allows the incorporation of diffraction into the calculation. Diffraction is topographically induced variations in wave direction. A model incorporating this complexity was developed by @isaeus2004, and is known as the "simplified wave model" (SWM). The model uses measurements of wind strength, fetch and empirically derived algorithims to mimic diffraction.

Numerical models are able to incorporate various physical process, as well as their complexity of interactions.

### 2018-08-28 proposal_draft_2018.Rmd
updated bib. file, added hyrodynamic modelling section

---
title: "Kelp morphometric properties and the link to hydrodynamic modelling"
author:
- affiliation: University of the Western Cape
  name: Ross Coppin, AJ Smit, Christo Rautenbach
date: "2018"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  word_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    toc: no
bibliography: mycollection2.bib   
---

@@

Marine ecosystems are maintained by a variety of complex interactions between abiotic and
biotic variables such as temperature, wave exposure, pH, competition, and processes such as top-down and bottom-up control, predator-prey relationships and phenology [@Doney2012]. These abiotic and biotic variables, the interactions between them, and the various ecological processes, ultimately determine the community composition and ecological functioning of all ecosystems. Climate directly and indirectly affects the way in which abiotic and biotic variables interact, but is often compounded by other impacts such as habitat destruction, pollution, and over-fishing [@Blamey2015]. Temperature and wave exposure have been recognised as important variables with regards to climate-driven changes within the ocean. In order to persist and survive within variable and changing environments, organisms must either migrate, adapt, or die.

>>>>

---
title: "Kelp morphometric properties and the link to hydrodynamic modelling"
author:
- affiliation: University of the Western Cape
  name: Ross Coppin, AJ Smit, Christo Rautenbach
date: "2018"
output:
  word_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    toc: no
  pdf_document:
    toc: yes
    toc_depth: 4
bibliography: mycollection.bib   
---

@@

Marine ecosystems are maintained by a variety of complex interactions between abiotic and
biotic variables such as temperature, wave exposure, pH, competition, and processes such as top-down and bottom-up control, predator-prey relationships and phenology [@Doney2012]. These abiotic and biotic variables, the interactions between them, and the various ecological processes, ultimately determine the community composition and ecological functioning of all ecosystems. Climate directly and indirectly affects the way in which abiotic and biotic variables interact, but is often compounded by other impacts such as habitat destruction, pollution, and over-fishing [@Blamey2015a]. Temperature and wave exposure have been recognised as important variables with regards to climate-driven changes within the ocean. In order to persist and survive within variable and changing environments, organisms must either migrate, adapt, or die.
